---
title: "Final Project Presentation"
author: "Julia Wilkins & Adrian Naranjo"
---

<br>

# Data

* Social Power NBA Dataset
* Found on Kaggle: https://www.kaggle.com/noahgift/social-power-nba
* Contains combined performance data for players in the 2016-2017 season, and salary, Twitter engagement, and Wikipedia traffic data. 

<br>

# Can we predict...

1. Salary from performance statistics?
2. Team valuations from individual salaries?
3. Salary based on social media stats or vice versa?
4. Points from other performance statistics?

<br>

# Exploratory Data Analysis

```{r, message=FALSE}
# Load Packages
library(tidyverse)
library(modelr)
library(boot)
library(leaps)
library(gridExtra)
library(gam)
```

```{r, message=FALSE, echo=FALSE, warning=FALSE}
# load some data for EDA

team_valuations <- read_csv('raw_data/nba_2017_team_valuations.csv') # Team vs. valuations

att_val_elo <- read_csv('raw_data/nba_2017_att_val_elo.csv') # Attendance, valuation, and ELO

salary <- read_csv('raw_data/nba_2017_salary.csv') %>% # 449 x 5
  mutate(SALARY2 = (SALARY / 1000000)) 


players_with_stats_combined <- read_csv('raw_data/nba_2017_players_stats_combined.csv') # 446 x 38 (PS)

players_with_salary_wiki_twitter <- read_csv('raw_data/nba_2017_players_with_salary_wiki_twitter.csv') # 239 x 42 (PS)
# Aggregates social media statistics at a team level
# Removes players which played for more than one team
team_twitter_wiki <- drop_na(players_with_salary_wiki_twitter) %>%
  group_by(TEAM) %>%
  summarise(total_pageviews = sum(PAGEVIEWS)
            ,total_twitter_favorite = sum(TWITTER_FAVORITE_COUNT)
            ,total_twitter_retweet = sum(TWITTER_RETWEET_COUNT)) %>%
  filter(!str_detect(TEAM, '/'))



```

<br>

### Exploring Team Data

<br>

```{r, message=FALSE, echo=FALSE}
# Simply visualizing the range of team values (combined salary)
# Maybe take this one out?
ggplot(team_valuations, aes(x = VALUE_MILLIONS
                            ,y = TEAM
                            ,size = VALUE_MILLIONS
                            ,color = VALUE_MILLIONS)) +
  geom_point() +
  labs(x = 'Team Value', y = 'Team Name') +
  ggtitle("Exploring Team Value")

# Shows attendance and team value are correlated
ggplot(att_val_elo, aes(x = VALUE_MILLIONS
                        ,y = AVG
                        ,size = VALUE_MILLIONS
                        ,color = VALUE_MILLIONS)) + 
  geom_smooth(color = 'brown') +
  geom_point() +
  labs(y = 'Attendance', x = 'Team Value') + 
  ggtitle("Team Value vs. Attendance ")
```

<br>

### Exploring Individual and Salary Data

<br>

```{r, message=FALSE, echo=FALSE}
# Display histogram of salaries... not surprisingly, right skewed
salary %>% 
  ggplot(aes(SALARY2, labels = TRUE)) +
  geom_histogram(binwidth = 2.5) +
  stat_bin(geom="text", aes(label=..count..), binwidth = 2.5) +
  labs(x = 'Salary in Millions', y = 'Number of Players') + 
  ggtitle("Individual Salary Distribution")
```
<br>
Quartiles of individual salaries:
```{r, message=FALSE}
# Find a quantiles
quantile(salary$SALARY, seq(0, 1, .1))
```
<br>
```{r, message=FALSE, echo=FALSE}

# Display salaries by position 
# Removed rows that only display a position a 'F' or 'G'
salary %>% 
  filter(POSITION %in% c('C', 'PF', 'PG', 'SF', 'SG')) %>% 
  ggplot(aes(POSITION, SALARY)) +
  geom_boxplot() +
  labs(x = 'Salary', y = 'Posiiton') + 
  ggtitle("Salary By Position")
```
<br>
Average salary by position:
```{r, message=FALSE, echo=FALSE}
# Calculate average salary/position
(avg_salary <- salary %>% 
    filter(POSITION %in% c('C', 'PF', 'PG', 'SF', 'SG')) %>% 
    group_by(POSITION) %>% 
    summarise(avg_salary = median(SALARY) / 1000000))



# Exploring salary vs. age
players_with_stats_salary <- inner_join(players_with_stats_combined, salary
                                        , by = c('PLAYER' =  'NAME'))
players_with_stats_salary <- select(players_with_stats_salary, -POSITION.y, -TEAM.y, -SALARY2)

players_with_stats_salary <- drop_na(players_with_stats_salary)

# Correlation between Age and Salary. Took out outliers.
players_with_stats_salary %>% 
  filter(AGE >= 20, AGE <= 35) %>% 
  ggplot(aes(x = AGE, y = SALARY)) +
  geom_point() +
  geom_smooth() + ggtitle("Age vs. Individual Salary")

```
<br><br>
```{r, message=FALSE, echo=FALSE}
# Correlation between between Player Effectiveness and Salary
ggplot(players_with_stats_salary, aes(x = PIE, y = SALARY)) +
  geom_point() +
  geom_smooth() + 
  ggtitle("Player Effectiveness vs. Salary")

```

<br>

### Social Media Stats by Team
```{r, message=FALSE, echo=FALSE}
team_social_media <- ggplot(team_twitter_wiki) + 
  geom_point(aes(total_pageviews, TEAM), color = 'red') +
  geom_point(aes(total_twitter_favorite, TEAM), color = 'blue') +
  geom_point(aes(total_twitter_retweet, TEAM), color = 'black') +
  labs(x = 'Social Media Statistics', y = 'Team') + 
  ggtitle("Social Media Stats per Team\n\n Red = WikiPageviews, Black = Twitter Retweets, Blue = Twitter Favorites")

team_social_media
```


<br>
<br>

# Main focus question: Can we predict salary from performance statistics? 

<br>